NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-03-31 paper/application.tex
finish full pass

In the previous section, we have noticed that a minor perturbation of the sentence may result in change of the final prediction (Fig.~\ref{fig:predictStability}(d)). Here, we want to make sense of what leads to the failed prediction. In this example, the premise \textbf{P} is ``A very young child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow'', and the original hypothesis \textbf{H1} is ``A child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow''. The perturbed hypothesis \textbf{H2} replaces the word \textbf{pile} with \textbf{heap} in \textbf{H1}. This example should be rather straightforward for the model since there are only minor differences between the \textbf{P} and \textbf{H1}/\textbf{H2}.

As illustrated in Fig.~\ref{fig:att2pred}(a), based on the graph attention visualization, we can see in the attention for (\textbf{P}, \textbf{H2}) pair that the words \textbf{pile} and \textbf{heap} are not well aligned (see Fig.~\ref{fig:att2pred}(a)).
%
To test whether the alignment is what contributes to the misclassification, the domain expert utilizes the attention editing functionality in the matrix attention view (Fig.~\ref{fig:attentionVis}(c)) to make the word \textbf{pile} align with \textbf{heap} (shown in Fig.~\ref{fig:att2pred}(b)).
%
As illustrated in Fig.~\ref{fig:att2pred} (d), after the edit, the original prediction (Fig.~\ref{fig:att2pred}(c)) has been moved from neutral to the classification boundary (Fig.~\ref{fig:att2pred}(d)), but the corrected attention fails to produce a conclusive \emph{entailment} prediction (a case where the attention edit corrects the final prediction is shown in Fig.~ref{fig:depTreeExample}).
%
%Interestingly, the original pair's attention is almost identical to the edited attention (the comparison is not shown in the figure). 
Such an observation implies that the model does not firmly believe ``\textbf{heap} of snow'' and ``\textbf{pile} of snow'' have the same meaning, which may indicate a potential issue with the encoder or word embedding.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scenario 3: Update the Model to Correct a Prediction}
%\subsection{Scenario 3: What Does It Take to Correct a Failed Prediction?}
Up to here, we have employed perturbation for input to understand prediction robustness, and utilized the perturbation of attention (and input) to infer the model decision-making process. Both these perturbation operations rely on forward propagation in the pipeline and assume the model parameter remains unchanged.
%
However, once we get a sense of how predictions are made and hypothesize about the cause of the failure in the case of a prediction error, it is natural to ask follow-up questions: What does it take to fix an incorrect prediction? And more importantly, what role does each of the three stages play in such a process? Are they affecting the prediction differently?

Domain experts can obtain answers to these questions by utilizing the prediction and pipeline view in the proposed tool. As discussed in detail in Section~\ref{sec:pipeline}, we employed a margin-infused relaxed algorithm (MIRA) based optimization with two objectives (apply the least amount of change to the parameter, and make the new prediction as close to the reassigned prediction as possible) to update the network parameters.
%
We then visualize how much each stage of the model is changed through the distribution of parameters' difference (see Fig.~\ref{fig:pipelineView}).

To infer the role each stage of the pipeline plays, the proposed tool allows the parameter update to be enabled or disabled for each pipeline stage.
The tool also includes an automatic option to test all the possible configuration combinations. As illustrated in Fig.~\ref{fig:pipelineUpdate}, the ground truth for this sentence pair is neutral. However, the model produces an incorrect label \emph{entailment}. The domain expert reassigns the prediction to \emph{neutral}, which triggers the prediction update optimization for $7$ different pipeline configurations. Four configurations are shown in Fig.~\ref{fig:pipelineUpdate}(a)(b)(c)(d). The updated predictions are illustrated as blue squares, and the pointers highlight the corresponding pipeline configurations.

@@

In the proposed system, we overlay the sentence dependency tree with the attention, which enables researchers to conduct comparisons between attention and grammar structure.
As illustrated in Fig.~\ref{fig:depTreeExample}(a), the prediction of the sentence pair (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is wrong. We can infer the cause by examining the attention, in which the word \textbf{green} in ``\textbf{green} jacket" is aligned to the \textbf{green} in ``\textbf{green} tea". Due to such an alignment, the model mistakenly believes the two \textbf{greens} are used to describe the same thing (therefore, predict \emph{entailment}).  However, as we examine the dependency tree, the two \textbf{greens} are attached to different words, i.e., ``green" in \textbf{P} is attached to ``jacket'', whereas ``green" in \textbf{H} is attached to ``tea''. Therefore, they should not be aligned to allow the classifier to make the right decision.
%
This experiment demonstrates the potential benefit of including grammar structure in the alignment computation. %This stresses the importance of dependency tree in our tool.
%
And interestingly, as illustrated in Fig.~\ref{fig:depTreeExample}(b), by editing the attention and forcing the alignment of the two \textbf{greens} to be zero, the prediction label is corrected (\emph{neutral}).

>>>>

In the previous section, we have noticed that a minor perturbation of the sentence may result in change of the final prediction (Fig.~\ref{fig:predictStability}(e)). Here, we want to make sense of what leads to the failed prediction. In this example, the premise \textbf{P} is ``A very young child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow'', and the original hypothesis \textbf{H1} is ``A child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow''. The perturbed hypothesis \textbf{H2} replaces the word \textbf{pile} with \textbf{heap} in \textbf{H1}. This example should be rather straightforward for the model since there are only minor differences between the \textbf{P} and \textbf{H1}/\textbf{H2}.

As illustrated in Fig.~\ref{fig:att2pred}(a), based on the graph attention visualization, we can see in the attention for (\textbf{P}, \textbf{H2}) pair that the words \textbf{pile} and \textbf{heap} are not well aligned.
%
To test whether the alignment is what contributes to the misclassification, the domain expert utilizes the attention editing functionality in the matrix attention view (Fig.~\ref{fig:attentionVis}(c)) to make the word \textbf{pile} align with \textbf{heap} (shown in Fig.~\ref{fig:att2pred}(b)).
%
After the edit, the original prediction (Fig.~\ref{fig:att2pred}(c)) has been moved from neutral to the classification boundary (Fig.~\ref{fig:att2pred}(d)), but the corrected attention fails to produce a conclusive \emph{entailment} prediction (another example, in which the attention edit corrects the final prediction, is shown in Fig.~\ref{fig:depTreeExample}).
%
%Interestingly, the original pair's attention is almost identical to the edited attention (the comparison is not shown in the figure). 
Such an observation implies that the model does not firmly believe ``\textbf{heap} of snow'' and ``\textbf{pile} of snow'' have the same meaning, which may indicate a potential issue with the encoder or word embedding.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scenario 3: Update the Model to Correct a Prediction}
%\subsection{Scenario 3: What Does It Take to Correct a Failed Prediction?}
Up to here, we have employed perturbation for input to understand prediction robustness, and utilized the perturbation of attention (and input) to infer the model decision-making process. Both these perturbation operations rely on forward propagation in the pipeline and assume the model parameter remains unchanged.
%
However, once we get a sense of how predictions are made and hypothesize about the cause of the failure in the case of a prediction error, it is natural to ask follow-up questions: What does it take to fix an incorrect prediction? And more importantly, what role does each of the three stages play in such a process? And, are they affecting the prediction differently?

Domain experts can obtain answers to these questions by utilizing the prediction and pipeline view in the proposed tool. As discussed in detail in Section~\ref{sec:pipeline}, we employed a margin-infused relaxed algorithm (MIRA) based optimization with two objectives (apply the least amount of change to the parameter, and make the new prediction as close to the reassigned prediction as possible) to update the network parameters.
%
We then visualize how much each stage of the model is changed through the distribution of differences between the two sets of parameters (see Fig.~\ref{fig:pipelineView}).

To infer the role each stage of the pipeline plays, the proposed tool allows the parameter update to be enabled or disabled for each pipeline stage.
The system also includes an automatic option to test all the possible configuration combinations. As illustrated in Fig.~\ref{fig:pipelineUpdate}, the ground truth for this sentence pair is \emph{neutral}. However, the model produces an incorrect label \emph{entailment}. The domain expert reassigns the prediction to \emph{neutral}, which triggers the prediction update optimization for $7$ different pipeline configurations. Four configurations are shown in Fig.~\ref{fig:pipelineUpdate}(a)(b)(c)(d). The updated predictions are illustrated as blue squares, and the arrowed lines highlight the corresponding pipeline configurations.

@@

In the proposed system, we overlay the sentence dependency tree with the attention, which enables researchers to conduct comparisons between attention and grammar structure.
As illustrated in Fig.~\ref{fig:depTreeExample}(a), the prediction of the sentence pair (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is wrong. We can infer the cause by examining the attention, in which the word \textbf{green} in ``\textbf{green} jacket" is aligned to the \textbf{green} in ``\textbf{green} tea". Due to such an alignment, the model mistakenly believes the two \textbf{greens} are used to describe the same thing (therefore, predict \emph{entailment}).  However, as we examine the dependency tree, the two \textbf{greens} are attached to different words, i.e., \textbf{green} in \textbf{P} is attached to ``jacket'', whereas \textbf{green} in \textbf{H} is attached to ``tea''. Therefore, they should not be aligned to allow the classifier to make the right decision.
%
This experiment demonstrates the potential benefit of including grammar structure in the alignment computation. %This stresses the importance of dependency tree in our tool.
%
And interestingly, as illustrated in Fig.~\ref{fig:depTreeExample}(b), by editing the attention and forcing the alignment of the two \textbf{greens} to be zero, the prediction label is corrected (\emph{neutral}).

### 2018-03-31 paper/discussion.tex
grammar fix
finish full pass

\section{Discussion}
The current setup for the proposed tool is suitable only for natural language inference task. However, due to the modular design and the many shared attributions among the end-two-end NLP models, we can readily extend the system to handle other tasks. In the future, we plan to open source our projection and add support for more NLP tasks, such as neural machine translation and question-and-answer.
%
From our evaluation process, we found that the quality of automatic perturbed sentences can be a potential limitation of the tool, since we rely on WordNet to generate the perturbation, which has a rather inclusive definition for synonymous. Often, we can identify perturbed sentences that are not particularly meaningful (e.g., sentences with very obscure words or usages). 
Even if all the words are meaningful, currently, there is no way we can verify whether the perturbed sentence is valid natural language composition or not.
%
However, from the natural language process point of view, the perturbation of a sentence while maintaining semantic and grammar correctness is an open research area on its own. 
%As better method appears, we can easily swap out the sentence perturbation functionality

To conclude, this work introduces a perturbation-driven visual interrogation system that provides experts with a streamlined exploration environment for testing hypothesis and obtaining intuition about the model. The proposed system frees the researchers from interruptions and tedious operations, therefore, allows them to focus on more productive activities.

>>>>

\section{Discussion}
The current setup for the proposed tool is suitable only for natural language inference task. However, due to the modular design and the many shared attributions among the end-to-end NLP models, we can readily extend the system to handle other tasks. In the future, we plan to open source our projection and add support for more NLP tasks, such as neural machine translation and question-and-answer.
%
From our evaluation process, we found that the quality of automatic perturbed sentences can be a potential limitation of the tool, since we rely on WordNet to generate the perturbation, which has a rather inclusive definition for synonymous. Often, we can identify perturbed sentences that are not particularly meaningful (e.g., sentences with very obscure words or usages). 
Even if all the words are meaningful, currently, there is no way we can verify whether the perturbed sentence is valid natural language composition or not.
%
However, from the natural language process point of view, the perturbation of a sentence while maintaining semantic and correct grammar is an open research area on its own. 
%As better method appears, we can easily swap out the sentence perturbation functionality

To conclude, this work introduces a perturbation-driven visual interrogation system that provides experts with a streamlined exploration environment for testing hypothesis and obtaining intuition about the model. The proposed system frees the researchers from interruptions and tedious operations, thereby allowing them to focus on more productive activities.

### 2018-03-31 paper/implementation.tex
finish full pass

%
In the proposed work, the NLI model is implemented in Python using \emph{pytorch}~\cite{PaszkeGrossChintala2017}.
The visual interface is implemented in Javascript using \emph{D3.js} library , and the Python server act as the glue between the Javascript visualization and the \emph{pytorch} model.

>>>>

%
In the proposed work, the NLI model is implemented in Python using \emph{pytorch}~\cite{PaszkeGrossChintala2017}.
The visual interface is implemented in Javascript using \emph{D3.js} library , and a Python server act as the glue between the Javascript visualization and the \emph{pytorch} model.
